---
title: "Literature_Review_visualizations"
author: "Libbie"
date: "2025-04-15"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r}
library(tidyverse)
```

```{r}
#install.packages("tidygeocoder")
library(tidygeocoder)

```

```{r}
library(readr)
data <- read_csv("C:\\Users\\User\\Desktop\\Literature Review\\coral_study_locations_for_plotting.csv")
class(data)
# Should say: "spec_tbl_df" "tbl_df" "tbl" "data.frame"

```

```{r}
glimpse(data)
view(data)
```
```{r}
data_coords <- data %>%
  geocode(Location, method = "osm", lat = latitude, long = longitude)

```
```{r}
head(data_coords)

```

```{r}
data_coords_clean <- data_coords %>%
  filter(!is.na(latitude) & !is.na(longitude))

```

```{r}
nrow(data_coords)         # Total number of rows before cleaning
nrow(data_coords_clean)   # Total number after removing NAs

```
```{r}
View(data_coords_clean)

```


```{r}
#install.packages("maps")
library(maps)

world_map <- map_data("world")

ggplot() +
  geom_polygon(data = world_map, aes(x = long, y = lat, group = group),
               fill = "gray95", color = "gray70") +
  geom_point(data = data_coords_clean, aes(x = longitude, y = latitude),
             color = "red", size = 2) +
  coord_fixed(1.3) +
  theme_minimal() +
  labs(
    title = "Mapped Coral Spawning Study Locations",
    x = "Longitude", y = "Latitude"
  )

```

```{r}
library(maps)

world_map <- map_data("world")

ggplot() +
  geom_polygon(data = world_map, aes(x = long, y = lat, group = group),
               fill = "gray95", color = "gray70") +
  geom_point(data = data_coords_clean, 
             aes(x = longitude, y = latitude, color = "Coral Spawning Studies from Current Databases"), 
             size = 2) +
  scale_color_manual(name = NULL, values = "red") +
  coord_fixed(1.3) +
  theme_minimal() +
  theme(legend.position = "bottom") +  # <- Move legend to the bottom
  labs(
    title = "Mapped Coral Spawning Study Locations",
    x = "Longitude", y = "Latitude"
  )

```

# NEED TO FIX INLAND POINT IN WA

```{r}
# View all geocoded coordinates
View(data_coords_clean)

```

```{r}
data_coords_clean %>%
  filter(longitude > 110, longitude < 125, latitude < -15, latitude > -35)
# isolating WA point
```

```{r}
data_coords_clean$latitude[which(data_coords_clean$Location == "Western Australia")] <- -22.0
data_coords_clean$longitude[which(data_coords_clean$Location == "Western Australia")] <- 113.9

```

```{r}
library(maps)

world_map <- map_data("world")

ggplot() +
  geom_polygon(data = world_map, aes(x = long, y = lat, group = group),
               fill = "gray95", color = "gray70") +
  geom_point(data = data_coords_clean, 
             aes(x = longitude, y = latitude, color = "Coral Spawning Studies from Current Databases"), 
             size = 2) +
  scale_color_manual(name = NULL, values = "red") +
  coord_fixed(1.3) +
  theme_minimal() +
  theme(legend.position = "bottom") +  # <- Move legend to the bottom
  labs(
    title = "Mapped Coral Spawning Study Locations",
    x = "Longitude", y = "Latitude"
  )
```



# Now we are going to add shaded areas that need more spawning data collection and attention 


```{r}
# Load required libraries
library(maps)       # For the world base map
library(ggplot2)    # For plotting
library(ggforce)    # For drawing ellipses
library(tibble)     # For tibble data format

# Load world map data
world_map <- map_data("world")

# Define underrepresented regions as large blue ellipses
gap_regions <- tibble::tibble(
  region = c(
    "Northern Indian Ocean", 
    "East Africa", 
    "Southeast Indian Ocean",
    "Central Pacific", 
    "High Latitude N. Pacific", 
    "Tropical Atlantic",
    "West Pacific", 
    "Caribbean", 
    "Southeast Asia"
  ),
  x0 = c(65, 45, 110, -170, 135, -38, 155, -75, 128),  # center longitudes
  y0 = c(10, -15, -12, 4, 30, -17, -10, 18, 5),        # center latitudes
  a = c(12, 10, 8, 15, 8, 6, 7, 15, 18),               # horizontal radii (width)
  b = c(6, 16, 6, 6, 4, 5, 5, 7, 12),                  # vertical radii (height)
  angle = 0                                            # no rotation
)

# Create the map
ggplot() +
  # Base world map polygons
  geom_polygon(data = world_map, aes(x = long, y = lat, group = group),
               fill = "gray95", color = "gray70") +
  
  # Red points for existing coral spawning studies
  geom_point(data = data_coords_clean, 
             aes(x = longitude, y = latitude, color = "Coral Spawning Studies from Current Databases"),
             size = 2) +
  
  # Large blue shaded ovals for under-monitored regions (with legend label)
  geom_ellipse(data = gap_regions,
               aes(x0 = x0, y0 = y0, a = a, b = b, angle = angle, fill = "Monitoring Gap"),
               alpha = 0.2, color = NA) +
  
  # Add legend entries for red and blue features
  scale_color_manual(name = NULL, values = "red") +
  scale_fill_manual(name = NULL, values = c("Monitoring Gap" = "blue")) +
  
  # Keep map proportions clean
  coord_fixed(1.3) +
  theme_minimal() +
  theme(legend.position = "bottom") +
  
  # Title and axis labels
  labs(
    title = "Coral Spawning Study Locations & Monitoring Gaps",
    x = "Longitude",
    y = "Latitude"
  )


```


# Now I will make my Environmental Spawning Cue Chart

```{r}
# Load required packages
library(tidyverse)

# Create a data frame of cue counts
cue_data <- tibble(
  Cue = c("Lunar Phase", "SST (Temp)", "Photoperiod / Light", "Rainfall", 
          "Tidal Cycles", "Salinity", "Wind", "Molecular"),
  Count = c(35, 30, 12, 5, 7, 3, 6, 4)
)

```

```{r}
# Plot cue coverage
ggplot(cue_data, aes(x = reorder(Cue, Count), y = Count)) +
  geom_col(fill = "mediumseagreen", width = 0.7) +
  coord_flip() +
  theme_minimal(base_size = 11) +
  labs(
    title = "Environmental Cues Reported in Current Databases",
    x = "Environmental Cue",
    y = "Number of Studies"
  ) +
  geom_text(aes(label = Count), hjust = -0.2) +
  theme(plot.title = element_text(face = "bold", hjust = 0.5))

```

# Now I am going to make a pie chart showcasing the species studied in my lit review

```{r}
library(tidyverse)

# Final pie chart data
taxonomic_data <- tibble(
  Taxa = c(
    "Acropora spp.", 
    "Porites spp.", 
    "Faviidae", 
    "Multiple or other brooding spp."
  ),
  Count = c(22, 6, 4, 12)
) %>%
  mutate(Percent = Count / sum(Count) * 100,
         Label = paste0(Taxa, " (", round(Percent), "%)"))

# Create the pie chart
ggplot(taxonomic_data, aes(x = "", y = Count, fill = Taxa)) +
  geom_col(width = 1, color = "white") +
  coord_polar(theta = "y") +
  theme_void() +
  labs(title = "Taxonomic Focus of Reviewed Coral Spawning Studies") +
  theme(
    plot.title = element_text(hjust = 0.5, face = "bold"),
    legend.title = element_blank()
  ) +
  geom_text(
    aes(label = paste0(round(Percent), "%")),
    position = position_stack(vjust = 0.5),
    color = "black",
    size = 5
  )

```
# Now I am going to make a Temporal Coverage Of Coral Spawning Studied in my Lit review

```{r}


# Create a dataset
temporal_data <- tibble(
  StudyDuration = c("One-time Observation", "Short-term (≤ 2 years)", "Long-term (> 2 years)"),
  Count = c(18, 17, 9)
) %>%
  mutate(Percent = Count / sum(Count) * 100)

# Convert to factor to control bar order (optional)
temporal_data$StudyDuration <- factor(temporal_data$StudyDuration, 
                                      levels = c("One-time Observation", 
                                                 "Short-term (≤ 2 years)", 
                                                 "Long-term (> 2 years)"))

# Plot 100% stacked bar chart
ggplot(temporal_data, aes(x = "", y = Percent, fill = StudyDuration)) +
  geom_bar(stat = "identity", width = 0.6, color = "white") +
  coord_flip() +
  theme_minimal(base_size = 14) +
  labs(
    title = "Temporal Resolution of Coral Spawning Studies",
    y = "Proportion of Studies (%)",
    x = NULL,
    fill = "Study Duration"
  ) +
  geom_text(aes(label = paste0(round(Percent), "%")),
            position = position_stack(vjust = 0.5),
            size = 5, color = "black") +
  scale_fill_brewer(palette = "Blues")

```
# trying a waffle chart for fun
```{r}
# Install the waffle package if needed
#install.packages("waffle")
library(waffle)
library(ggplot2)

# Create named vector of study durations
temporal_data <- c(
  `One-time Observation` = 18,
  `Short-term (≤ 2 years)` = 17,
  `Long-term (> 2 years)` = 9
)

# Plot the waffle chart
waffle(
  parts = temporal_data,
  rows = 5,  # 5 rows of boxes → 44 total boxes
  size = 1,
  colors = c("#4B9CD3", "#74C476", "#FDD835"),
  title = "Temporal Resolution of Coral Spawning Studies",
  legend_pos = "bottom"
)

```

# trying a horizontal bar chart

```{r}
#library(tidyverse)

# Data
temporal_data <- tibble(
  Duration = c("One-time Observation", "Short-term (≤ 2 years)", "Long-term (> 2 years)"),
  Count = c(18, 17, 9)
) %>%
  mutate(Percent = Count / sum(Count) * 100,
         Label = paste0(Duration, "\n", round(Percent), "%"))

# Reorder for stacking
temporal_data$Duration <- factor(temporal_data$Duration,
                                  levels = c("Long-term (> 2 years)", "Short-term (≤ 2 years)", "One-time Observation"))

# Final plot
ggplot(temporal_data, aes(x = "", y = Percent, fill = Duration)) +
  geom_bar(stat = "identity", width = 0.6, color = "white") +
  coord_flip() +
  geom_text(aes(label = Label), position = position_stack(vjust = 0.5), size = 4.5) +
  theme_void() +
  scale_fill_manual(values = c("#FDD835", "#74C476", "#4B9CD3")) +
  labs(title = "Temporal Resolution of Coral Spawning Studies") +
  theme(
    plot.title = element_text(
      hjust = 0.5,
      size = 14,
      face = "bold",
      margin = margin(b = 2)  # minimal space below title
    ),
    plot.margin = margin(t = 5, r = 10, b = 10, l = 10),  # tighter top margin
    legend.position = "none"
  )


```




